---
title: "OECD TiVa LMIC GVCs"
author: "Victor Kummritz and Bastiaan Quast"
date: "March 24, 2015"
output: ioslides_presentation
---

## Motivation

 - Global Value Chains relevance
 - GVCs in developement (capacity building)
 - General GVC picture overemphasises north-north because of higher monetary value

## This paper

## Data

ICIOs from TiVa (not WIOD)

TiVa data
 - 57 countries
 - 34 sectors
 - LMICs

## New Toolbox and Indicators

 - source / using countries specific indicators
 - decompor and gvc functions
 
 ```{r}
 library(decompr)
 ls.str("package:decompr")
 ```
 
 ```{r}
 library(gvc)
 ls.str("package:gvc")
 ```

## General Stylised Facts

 - section 3.1 of wwz
 - growth of GVC trade
 - length of GVC
 - determinants of an industry's participation
 - determinants of a country's participation
 - trade balanses in VA terms differ substantially from those in gross terms


## GVCs of LMICs

 - stylised facts for LMICs
 - source / using countries specific indicators for LMICs
 - nrca and trca of LMIC
 - fvafin vs. fvaint for LMICs (WWZ 3.3)
 - dvix decomposition (WWZ 3.4)

## Conclusion
